---
title: "Twitter's Birdwatch Platform: Replicating Figures from the Original Paper"
runtime: shiny
author: "Alou Kone and Vaishnavi Dixit"
date: "June 17, 2025"
output:  
  html_document:
    code_folding: show
---
We attempted to replicate the figures from the original paper on Twitter's Birdwatch platform. The original paper can be found [here](https://arxiv.org/pdf/2104.07175).
Some of the aspects of the figures were not replicated due to the lack of access to the original data. However, we were able to replicate most of the figures using the data we had access to. The figures are presented below.

Original Datasets Size: 11,802 Birdwatch notes and 52,981 ratings
Our Datasets Dize: 9,377 Birdwatch notes and 45,885 ratings

### After Importing Required Libraries and Data (Notes, Ratings and Source Tweets)
```{r setup, include=FALSE}

library(tidyverse)
library(scales)
library(modelr)
library(readr)
library(ggplot2)
library(dplyr)
library(forcats)

notes <- read_tsv("C:/Users/ds3/Desktop/New folder/coursework/week4/community-notes-2025-group-6/notes.tsv")
ratings <- read_tsv("C:/Users/ds3/Desktop/New folder/coursework/week4/community-notes-2025-group-6/complete_filtered_ratings.tsv")

load("C:/Users/ds3/Desktop/New folder/coursework/week4/community-notes-2025-group-6/source_tweets.Rdata", ournew_env <- new.env())
source_tweets <- ournew_env[["."]]
source_tweets <- source_tweets %>% filter(!is.na(source_followers_count))
nrow(notes)
nrow(ratings)
```

### Figure 2: Distribution of Birdwatch Notes by Classification and Trustworthiness
```{r, fig.width=6, fig.height=3, echo=FALSE}

notes %>% mutate(classification = ifelse(classification == "MISINFORMED_OR_POTENTIALLY_MISLEADING", "Misleading", "Not Misleading")) %>%
  mutate(trustworthiness = ifelse(trustworthySources == 1, "Trustworthy Sources", "No Trustworthy Sources")) %>%
  ggplot(aes(y = classification, fill = trustworthiness)) +
  geom_bar() +
  scale_fill_manual(
    values = c("No Trustworthy Sources" = "yellow", "Trustworthy Sources" = "#4141db")) +
  labs(x = "Count", y = "Classification", fill = "Trustworthiness") +
  theme_minimal() +
  theme(legend.position = "top") 
```

### Figure 3: Types of Misleading Birdwatch Notes
```{r, fig.width=6, fig.height=3, echo=FALSE}

notes_misleading <- notes %>% rename(Other = misleadingOther,  Factual_Error = misleadingFactualError, Missing_Important_Context = misleadingMissingImportantContext, 
  Unverified_Claim_As_Fact = misleadingUnverifiedClaimAsFact, Outdated_Information = misleadingOutdatedInformation, 
  Satire = misleadingSatire, Manipulated_Media = misleadingManipulatedMedia) %>% 
  summarize(Other = sum(Other, na.rm = TRUE), Factual_Error = sum(Factual_Error, na.rm = TRUE), Manipulated_Media = sum(Manipulated_Media, na.rm = TRUE), 
  Outdated_Information = sum(Outdated_Information, na.rm = TRUE), Missing_Important_Context = sum(Missing_Important_Context, na.rm = TRUE), 
  Unverified_Claim_As_Fact = sum(Unverified_Claim_As_Fact, na.rm = TRUE), 
  Satire = sum(Satire, na.rm = TRUE))

notes_misleading <- pivot_longer(notes_misleading, cols = c(Other, Factual_Error, Manipulated_Media, Outdated_Information, Missing_Important_Context, Unverified_Claim_As_Fact, 
  Satire), names_to = "misleadingType", values_to = "count") %>% arrange(desc(count))


ggplot(notes_misleading, aes(x = count, y = fct_reorder(misleadingType, count))) +
  geom_bar(stat = "identity", fill = "red") +
  theme_minimal() +
  labs(x = "Number of Birdwatch Notes", y = "Misleading Type")
```

### Figure 4: Types of Not Misleading Birdwatch Notes
```{r, fig.width=6, fig.height=3, echo=FALSE}

notes_not_misleading <- notes %>% rename(Other = notMisleadingOther,  Factually_Correct = notMisleadingFactuallyCorrect, Personal_Opinion = notMisleadingPersonalOpinion, 
  Outdated_But_Not_When_Written = notMisleadingOutdatedButNotWhenWritten, Clearly_Satire = notMisleadingClearlySatire) %>% 
  summarize(Other = sum(Other, na.rm = TRUE), Factually_Correct = sum(Factually_Correct, na.rm = TRUE), Personal_Opinion = sum(Personal_Opinion, na.rm = TRUE), 
  Clearly_Satire = sum(Clearly_Satire, na.rm = TRUE), Outdated_But_Not_When_Written = sum(Outdated_But_Not_When_Written, na.rm = TRUE))

notes_not_misleading <- pivot_longer(notes_not_misleading, cols = c(Other, Factually_Correct, Personal_Opinion, Clearly_Satire, Outdated_But_Not_When_Written), 
  names_to = "misleadingType", values_to = "count") %>% arrange(desc(count))

ggplot(notes_not_misleading, aes(x = count, y = fct_reorder(misleadingType, count))) +
  geom_bar(stat = "identity", fill = "blue") +
  theme_minimal() +
  labs(x = "Number of Birdwatch Notes", y = "Not Misleading Type")
```

### Figure 5c: Distribution of Birdwatch Notes by Classification and Word Count
```{r, fig.width=6, fig.height=5, echo=FALSE, warning=FALSE}
notes <- notes %>% group_by(classification) %>%
  mutate(word_count = str_count(summary, boundary("word"))) %>% mutate(total_words = sum(word_count, na.rm = TRUE))

notes %>%
  group_by(classification) %>%
  arrange(word_count) %>%
  mutate(
    total = n(),
    ccdf = (1 - (row_number()) / total)*100
  ) %>%
  ggplot(aes(x = word_count, y = ccdf, color = classification)) +
  geom_line() +
  theme_minimal() +
  scale_y_log10(limits = c(0.01, 100), label = comma)+
  theme(legend.position = "top") +
  labs(x = "Word Count", y = "CCDF (%)")
```

### Figure 7a: Distribution of Birdwatch Notes by Classification and Helpful Ratio
```{r, fig.width=6, fig.height=5, echo=FALSE}
# Calculate ratio_helpful for each noteId
ratings_summary <- ratings %>%
  group_by(noteId) %>% 
  mutate(
    helpful = ifelse(
      version == 1,
      helpful,
      ifelse(helpfulnessLevel %in% c("HELPFUL", "SOMEWHAT_HELPFUL"), 1, 0)
    )
  ) %>%
  summarize(
    total_helpful = sum(helpful == 1, na.rm = TRUE),
    total_unhelpful = sum(helpful == 0, na.rm = TRUE)
  ) %>%
  mutate(
    total = total_helpful + total_unhelpful,
    ratio_helpful = ifelse(total == 0, NA, total_helpful / total)
  )

# Join with notes and calculate CCDF by classification
notes_ratings <- notes %>%
  inner_join(ratings_summary, by = "noteId") %>%
  group_by(classification) %>%
  arrange(ratio_helpful) %>%
  mutate(
    rank = min_rank(ratio_helpful),
    n = n(),
    ccdf = (1 - ((rank-1) / n)) * 100
  )

# Plot
notes_ratings %>%
  ggplot(aes(x = ratio_helpful, y = ccdf, color = classification)) +
  geom_line() + 
  theme(legend.position = "top") +
  labs(x = "Helpful Ratio", y = "CCDF (%)")
```

### Figure 7b: Distribution of Votes by Classification
```{r, fig.width=6, fig.height=5, echo=FALSE}
# Calculate ratio_helpful for each noteId
ratings_summary_votes <- ratings_summary

# Join with notes and calculate CCDF by classification
notes_ratings_votes <- notes %>%
  inner_join(ratings_summary_votes, by = "noteId") %>%
  group_by(classification) %>%
  arrange(total) %>%
  mutate(
    rank = min_rank(total),
    n = n(),
    ccdf = (1 - ((rank-1) / n)) * 100
  )

# Plot
notes_ratings_votes %>%
  ggplot(aes(x = total, y = ccdf, color = classification)) +
  geom_line()+
  theme(legend.position = "top") +
  scale_y_log10(limits = c(0.01, 100), label = comma)+
  labs(x = "Votes (Helpful & not Helpful)", y = "CCDF (%)")
```

### Figure 8: Distribution of Helpful Ratings
```{r, fig.width=6, fig.height=3, echo=FALSE}

ratings_helpful <- ratings %>% rename(Other = helpfulOther,  Informative = helpfulInformative, Clear = helpfulClear, Empathetic = helpfulEmpathetic, 
  Good_Sources = helpfulGoodSources, Unique_Context = helpfulUniqueContext, Address_Claim = helpfulAddressesClaim, Important_Context = helpfulImportantContext) %>% 
  summarize(Other = sum(Other, na.rm = TRUE), Informative = sum(Informative, na.rm = TRUE), Clear = sum(Clear, na.rm = TRUE), Empathetic = sum(Empathetic, na.rm = TRUE), 
  Good_Sources = sum(Good_Sources, na.rm = TRUE), Unique_Context = sum(Unique_Context, na.rm = TRUE), Address_Claim = sum(Address_Claim, na.rm = TRUE), 
  Important_Context = sum(Important_Context, na.rm = TRUE))

ratings_helpful <- pivot_longer(ratings_helpful, cols = c(Other, Informative, Clear, Empathetic, Good_Sources, Unique_Context, Address_Claim, Important_Context), 
  names_to = "helpingType", values_to = "count") %>% arrange(desc(count))

ggplot(ratings_helpful, aes(x = count, y = fct_reorder(helpingType, count))) +
  geom_bar(stat = "identity", fill = "blue") +
  theme_minimal() +
  labs(x = "Number of Ratings", y = "Helping Type")
```

### Figure 9: Distribution of Not Helpful Ratings
```{r, fig.width=6, fig.height=3, echo=FALSE}
ratings_unhelpful <- ratings %>% rename(Other = notHelpfulOther,  Incorrect = notHelpfulIncorrect, Sources_Missing_Or_Unreliable = notHelpfulSourcesMissingOrUnreliable,  
  Opinion_Speculation_Or_Bias = notHelpfulOpinionSpeculationOrBias, Missing_Key_Points = notHelpfulMissingKeyPoints, Outdated = notHelpfulOutdated, 
  Hard_To_Understand = notHelpfulHardToUnderstand, Argumentative_Or_Biased = notHelpfulArgumentativeOrBiased, Off_Topic = notHelpfulOffTopic, 
  Spam_Harassment_Or_Abuse = notHelpfulSpamHarassmentOrAbuse, Irrelevant_Sources = notHelpfulIrrelevantSources) %>% summarize(Other = sum(Other, na.rm = TRUE), 
  Incorrect = sum(Incorrect, na.rm = TRUE), Sources_Missing_Or_Unreliable = sum(Sources_Missing_Or_Unreliable, na.rm = TRUE), 
  Opinion_Speculation_Or_Bias = sum(Opinion_Speculation_Or_Bias, na.rm = TRUE), Missing_Key_Points = sum(Missing_Key_Points, na.rm = TRUE), 
  Outdated = sum(Outdated, na.rm = TRUE), Hard_To_Understand = sum(Hard_To_Understand, na.rm = TRUE), Argumentative_Or_Biased = sum(Argumentative_Or_Biased, na.rm = TRUE), 
  Off_Topic = sum(Off_Topic, na.rm = TRUE), Spam_Harassment_Or_Abuse = sum(Spam_Harassment_Or_Abuse, na.rm = TRUE), Irrelevant_Sources = sum(Irrelevant_Sources, na.rm = TRUE))

ratings_unhelpful <- pivot_longer(ratings_unhelpful, cols = c(Other, Incorrect, Sources_Missing_Or_Unreliable,  Opinion_Speculation_Or_Bias, Missing_Key_Points, Outdated, 
  Hard_To_Understand, Argumentative_Or_Biased, Off_Topic, Spam_Harassment_Or_Abuse, Irrelevant_Sources), names_to = "nothelpingType", values_to = "count") %>% 
  arrange(desc(count))

ggplot(ratings_unhelpful, aes(x = count, y = fct_reorder(nothelpingType, count))) +
  geom_bar(stat = "identity", fill = "red") +
  theme_minimal() +
  labs(x = "Number of Ratings", y = "Not Helping Type")
```

### Figure 10: Regression Analysis of Helpful Ratings
```{r , fig.width=6, fig.height=5, echo=FALSE, warning=FALSE}
notes_ratings_reg <- inner_join(notes, source_tweets, by = c("noteId")) %>% inner_join(ratings, by = "noteId") 
len <- nrow(notes_ratings_reg)

notes_ratings_reg <- mutate(notes_ratings_reg, classification = ifelse(classification == "MISINFORMED_OR_POTENTIALLY_MISLEADING", 1, 0), 
  age = as.integer(interval(source_account_created_at, Sys.Date()) / years(1)), word_count = str_count(summary, boundary("word")))


notes_ratings_reg <- mutate(notes_ratings_reg, word_count = scale(word_count), source_followers_count = scale(source_followers_count), 
 source_friends_count = scale(source_friends_count), age = scale(age))


reg_model <- glm(helpful ~ classification + trustworthySources + word_count + source_followers_count + source_verified + source_friends_count + 
  age, family = binomial, data = notes_ratings_reg)


coefficients_ratio <- as.data.frame(summary(reg_model)$coefficients)

coefficients_ratio$predictor <- rownames(coefficients_ratio)

ggplot(coefficients_ratio) + geom_point(aes(x = predictor, y = Estimate), color = "blue", size = 3) + 
  geom_errorbar(aes(x = predictor, ymin = Estimate - (2.576 * coefficients_ratio$"Std. Error"), 
  ymax = Estimate + (2.576 * coefficients_ratio$"Std. Error")), width = 0, color = "orange") +
  labs(x = "Predictor Variables", y = "Coefficient Estimate") +
  theme_minimal() + 
  scale_y_continuous(limits = c(-0.8, 1.0)) + 
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

```

Research Question:
Can we see a connection between followers and followees count and if it affects the helpfulness 
of a note as a famous person can have more followers and spread wrong information and if a person 
who has less stats but their correct info can be rated less helpful?

For the extension we can check the relationship between notes classified as notmisleading and helpfulness.
That is we can check the pearson r cooleration beetween notes classified as notmisleading and the feautre helpful.
We could also use a decissionTree algorthim with feautre interction beetween the different types of (missleading, notmisleading) and the of
(helpful, nothelpful). To see how different notes groups, this might encorver another layer of complexity where its not so balck and white, 
and we might choose to show notes classified as misleading instead.